---
title: "2018-02-08 Pan-NF Drug Screening TIMMA Combination Prediction"
author: "Robert Allaway"
date: "2/8/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(timma)
library(synapser)
synLogin()

targets <- synGet("syn17091507")$path %>% readRDS() %>% filter(mean_pchembl > 6)

```
##Prepare the drug-target dataset. 

We've already done the hard work of mapping these two datasets together, so we just need to filter down to get a binary list of targets. We'll assume everything with a pChembl >6 is a "real" target.

```{r}

targets <- synGet("syn17091507")$path %>% readRDS() %>% 
  filter(mean_pchembl > 6) %>% 
  select(internal_id, hugo_gene, std_name) %>% 
  distinct()

```
## Prepare the drug screening data. 

This data is standardized across all public NF studies, and a variety of drug metrics were calculated each experiment. However, there is a _lot_ of diverse data here, so we should do a bit to clean up the data and narrow the scope, and eliminate metrics that might not be that accurate. 

Specificially, we will: 
-get the dataset 
-filter for absolute IC50 values only, also known as GI50 (this different than the relative IC50, which is the half-maximal inhibition)
-filter for one cell line model - this method can only work on set of drugs on a single model. Let's look at the NF1 deficient pNF cell line ipNF95.6.
-filter out the combination data (too messy for a first pass of this method)
-select only the columns we need ()
-`inner_join` on the target dataset to filter out any drugs that we don't have target info for. 

```{r}

drug_data <- synGet("syn17462699")$path %>% read.csv() 

drug_data_filt <- drug_data %>% 
  filter(response_type == "IC50_abs") %>% 
  filter(model_name == "ipNF95.6") %>% 
  group_by(drug_screen_id) %>% 
  filter(n() == 1) %>% 
  ungroup() %>% 
  select(DT_explorer_internal_id, drug_name, response) %>% 
  inner_join(targets, by = c("DT_explorer_internal_id" = "internal_id"))
  
target_data_filt <- drug_data_filt %>% 
  select(DT_explorer_internal_id, hugo_gene) %>% 
  distinct() %>% 
  add_column(foo = c(1)) %>% 
  spread(DT_explorer_internal_id, foo, fill = 0)



```


```{r , echo=FALSE}

```
