---
title: "2018-02-08 pNF Drug Screening TIMMA Combination Prediction"
author: "Robert Allaway"
date: "2/8/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(timma)
library(synapser)
synLogin()

```
##Prepare the drug-target dataset. 

We've already done the hard work of mapping these two datasets together, so we just need to filter down to get a binary list of targets. We'll assume everything with a pChembl >6 is a "real" target.

```{r}

targets <- synGet("syn17091507")$path %>% readRDS() %>% 
  filter(mean_pchembl > 7) %>% 
  select(internal_id, hugo_gene, std_name) %>% 
  distinct()

```
## Prepare the drug screening data. 

This data is standardized across all public NF studies, and a variety of drug metrics were calculated each experiment. However, there is a _lot_ of diverse data here, so we should do a bit to clean up the data and narrow the scope, and eliminate metrics that might not be that accurate. 

Specificially, we will: 
-get the dataset 
-filter for absolute IC50 values only, also known as GI50 (this different than the relative IC50, which is the half-maximal inhibition)
-filter for NF1 deficient pNF cell lines
-remove outlier cell line
-filter out the combination data (too complicated for a first pass of this method)
-calculate the median IC50 for each drug across all cell lines
-select only the information we need
-`inner_join` on the target dataset to filter out any drugs that we don't have target info for. 


```{r}

drug_data <- synGet("syn17462699")$path %>% read.csv() 

pnf <- c("ipNF05.5", "ipNF06.2A", "ipNF95.11b C/T", "ipnNF95.11C", "ipNF95.6", "ipNF05.5 (mixed clone)", "ipNF95.11b C")

drug_data_filt <- drug_data %>% 
  filter(response_type == "IC50_rel") %>% 
  filter(model_name %in% pnf) %>% 
  group_by(drug_screen_id) %>% 
  filter(n() == 1) %>% 
  ungroup() %>%  
  mutate(response = case_when(response < 50 ~ response,
                              response >= 50 ~ 50))

ggplot(data = drug_data_filt) +
  geom_boxplot(aes(x = model_name, y = response))
  
drug_data_filt <- drug_data_filt %>% 
  filter(model_name != "ipNF95.11b C") %>% 
  group_by(DT_explorer_internal_id) %>% 
  filter(n() > 3) %>% 
  ungroup() %>% 
  select(DT_explorer_internal_id, response) %>%
  group_by(DT_explorer_internal_id) %>% 
  summarize('median_response' = median(response))%>% 
  ungroup() %>% 
  inner_join(targets, by = c("DT_explorer_internal_id" = "internal_id"))
  
target_data_filt <- drug_data_filt %>% 
  select(DT_explorer_internal_id, hugo_gene) %>% 
  distinct() %>% 
  add_column(foo = c(1)) %>% 
  spread(hugo_gene, foo, fill = 0) %>% 
  column_to_rownames("DT_explorer_internal_id") %>% 
  as.matrix()

drug_data_filt_distinct <- drug_data_filt %>% 
  select(DT_explorer_internal_id, median_response) %>% 
  distinct() %>% 
  mutate(scaled_response = BBmisc::normalize(median_response, method = "range", range = c(1,0), margin = 1L))

```


```{r , echo=FALSE}
output <- timma(x = target_data_filt, y = drug_data_filt_distinct$scaled_response, sp = 1,
max_k = 8, filtering = FALSE , class = 2, averaging = "one.sided",
weighted = FALSE, verbosity = T)
```

```{r}
perplexities <- c(5,10,30,50,70,100)

for(i in perplexities){
foo <- Rtsne(target_data_filt, check_duplicates = F, perplexity = i)

foo2 <- foo$Y %>% 
  as.data.frame() %>% 
  add_column("DT_explorer_internal_id" = as.integer(rownames(target_data_filt))) %>% 
  left_join(drug_data_filt) %>% 
  left_join(distinct(select(targets, internal_id, std_name)), by = c("DT_explorer_internal_id" = "internal_id"))

p<-ggplot(foo2) + 
  geom_point(aes(x = V1, y = V2), alpha = 0.5)

print(p)
}
```

```{r}

foo <- Rtsne(target_data_filt, check_duplicates = F, perplexity = 30)

foo2 <- foo$Y %>% as.data.frame() %>% add_column("DT_explorer_internal_id" = as.integer(rownames(target_data_filt))) %>% left_join(drug_data_filt) %>% left_join(distinct(select(targets, internal_id, std_name)), by = c("DT_explorer_internal_id" = "internal_id"))

ggplot(foo2) + geom_point(aes(x = V1, y = V2)) + scale_color_viridis_c(option = "C")  

ggplot(foo2) + geom_point(aes(x = V1, y = V2, color = log(median_response)), alpha = 0.5) + scale_color_viridis_c(option = "C")  

ggplot(foo2) + geom_point(aes(x = V1, y = V2, color = log(median_response)), alpha = 0.5) + scale_color_viridis_c(option = "C")  + ggrepel::geom_label_repel(data = foo2 %>% filter(median_response < 0.05), aes(x = V1, y = V2, label = std_name), size = 2)

ggplot(foo2) + geom_point(aes(x = V1, y = V2, color = log(median_response)), alpha = 0.5) + scale_color_viridis_c(option = "C")  + ggrepel::geom_label_repel(data = foo2 %>% filter(median_response < 0.15), aes(x = V1, y = V2, label = std_name), size = 2)
```

